### 2014-11-12 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Minor edits to references to FDS throughout

% PDF metadata
\hypersetup{
  pdftitle={Simulation of a Fire in a Hillside Residential Structure - San Francisco, CA},
  pdfauthor={Overholt, Kristopher J., Weinschenk, Craig G., and Madrzykowski, Daniel},
  pdfsubject={The Fire Dynamics Simulator (FDS) fire model, which is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on June 2, 2011, within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the rear basement windows began to fail, the interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure and vent openings located on the front side of the structure. The temperature of the gases in the interior stairwell was estimated to be in excess of 700~C (1300~F). Two firefighters were located in the flow path between the basement and the doors on the front side of the structure, were exposed to these elevated temperatures, and later died as a result of their injuries.},
  pdfkeywords={flow path; fire dynamics; LODD; fire modeling; FDS simulation}
}

@@

The Fire Dynamics Simulator (FDS) fire model, which is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on June 2, 2011, within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the rear basement windows began to fail, the interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure and vent openings located on the front side of the structure. The temperature of the gases in the interior stairwell was estimated to be in excess of 700~$^{\circ}$C (1300~$^{\circ}$F). Two firefighters were located in the flow path between the basement and the doors on the front side of the structure, were exposed to these elevated temperatures, and later died as a result of their injuries.

@@

Part of the function of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to develop and apply technology, measurements, and standards to improve the understanding of the behavior, prevention, and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) fire model to provide insight into the fire development and thermal conditions of fires that have resulted in line of duty deaths (LODDs)~\cite{Madrzykowski:1,Iowa,Texas,Bryner:Charleston,barowy:texas,Weinschenk:Chicago}. The objective of these studies has been to improve the safety and effectiveness of firefighters.

On June 2, 2011, a fire in a multi-level, wood-frame residential structure claimed the life of two firefighters of the San Francisco Fire Department. NIST examined the fire dynamics of this incident at the request of the National Institute for Occupational Safety and Health (NIOSH) and the San Francisco Fire Department. A computer simulation of the fire incident was conducted using the NIST Fire Dynamics Simulator~\cite{FDS_Users_Guide} and Smokeview~\cite{Smokeview_Users_Guide} software to provide insight into the fire development and thermal conditions that likely existed in the residence during the fire. The specific objectives of the simulation detailed in this report are:
\begin{enumerate}
\item To examine the effect of fire-induced flow paths (including temperature, pressure, and fire conditions) in this multi-level residential structure using physics-based calculations.
\item To provide visualizations of the fire behavior that are representative of the conditions that members of the San Francisco Fire Department likely experienced during the course of their interior operations.
\end{enumerate}
This document describes the input and the results of the FDS (version 6.1.2) simulation. The simulation was developed using a combination of knowledge of the fire scenario and appropriate engineering approximations and assumptions. Analysis of the simulation results focuses on the hazardous conditions that developed following the development of a flow path inside of the structure. This document is organized as follows: Section~\ref{sec:incident_summary} provides a summary of the fire incident, Section~\ref{sec:model_description} describes the relevant model input parameters and assumptions that were used to develop the simulation, Section~\ref{sec:model_results} presents the simulation results, and Section~\ref{sec:discussion} discusses the simulation results as they relate to firefighter safety and effectiveness. Appendix~\ref{sec:drawings} contains dimensioned drawings of the basement, first, and second floors of the structure.

@@

NIST's Fire Dynamics Simulator was used to provide insight into the fire dynamics of a fire that occurred within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The fuel, fire size, and fire growth rate that were used in the FDS simulation were estimated by taking into account all of the available information including the NIOSH report~\cite{NIOSH:Bowyer2}, post-incident pictures, and relevant literature. This resulted in a maximum specified source fire of approximately 20~MW in the basement and rear balcony of the structure. Based on the limited ventilation conditions in the basement, the FDS simulation results indicated a HRR of approximately 2~MW as a result of the limited supply of oxygen. After the rear basement windows failed, the HRR increased to approximately 32~MW before it reached a steady-state value of approximately 20~MW.

>>>>

% PDF metadata
\hypersetup{
  pdftitle={Simulation of a Fire in a Hillside Residential Structure - San Francisco, CA},
  pdfauthor={Overholt, Kristopher J., Weinschenk, Craig G., and Madrzykowski, Daniel},
  pdfsubject={Fire Dynamics Simulator (FDS), which is a fire model that is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on June 2, 2011, within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the rear basement windows began to fail, the interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure and vent openings located on the front side of the structure. The temperature of the gases in the interior stairwell was estimated to be in excess of 700~C (1300~F). Two firefighters were located in the flow path between the basement and the doors on the front side of the structure, were exposed to these elevated temperatures, and later died as a result of their injuries.},
  pdfkeywords={flow path; fire dynamics; LODD; fire modeling; FDS simulation}
}

@@

Fire Dynamics Simulator (FDS), which is a fire model that is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on June 2, 2011, within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the rear basement windows began to fail, the interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure and vent openings located on the front side of the structure. The temperature of the gases in the interior stairwell was estimated to be in excess of 700~$^{\circ}$C (1300~$^{\circ}$F). Two firefighters were located in the flow path between the basement and the doors on the front side of the structure, were exposed to these elevated temperatures, and later died as a result of their injuries.

@@

Part of the function of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to develop and apply technology, measurements, and standards to improve the understanding of the behavior, prevention, and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used Fire Dynamics Simulator (FDS) to provide insight into the fire development and thermal conditions of fires that have resulted in line of duty deaths (LODDs)~\cite{Madrzykowski:1,Iowa,Texas,Bryner:Charleston,barowy:texas,Weinschenk:Chicago}. The objective of these studies has been to improve the safety and effectiveness of firefighters.

On June 2, 2011, a fire in a multi-level, wood-frame residential structure claimed the life of two firefighters of the San Francisco Fire Department. NIST examined the fire dynamics of this incident at the request of the National Institute for Occupational Safety and Health (NIOSH) and the San Francisco Fire Department. A computer simulation of the fire incident was conducted using Fire Dynamics Simulator~\cite{FDS_Users_Guide} and Smokeview~\cite{Smokeview_Users_Guide} to provide insight into the fire development and thermal conditions that likely existed in the residence during the fire. The specific objectives of the simulation detailed in this report are:
\begin{enumerate}
\item To examine the effect of fire-induced flow paths (including temperature, pressure, and fire conditions) in this multi-level residential structure using physics-based calculations.
\item To provide visualizations of the fire behavior that are representative of the conditions that members of the San Francisco Fire Department likely experienced during the course of their interior operations.
\end{enumerate}
This document describes the input and the results of the FDS (version 6.1.2) simulation. The simulation was developed using a combination of knowledge of the fire scenario and appropriate engineering approximations and assumptions. Analysis of the simulation results focuses on the hazardous conditions that developed following the development of a flow path inside of the structure. This document is organized as follows: Section~\ref{sec:incident_summary} provides a summary of the fire incident, Section~\ref{sec:model_description} describes the relevant model input parameters and assumptions that were used to develop the simulation, Section~\ref{sec:model_results} presents the simulation results, and Section~\ref{sec:discussion} discusses the simulation results as they relate to firefighter safety and effectiveness. Appendix~\ref{sec:drawings} contains dimensioned drawings of the basement, first, and second floors of the structure.

@@

Fire Dynamics Simulator was used to provide insight into the fire dynamics of a fire that occurred within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The fuel, fire size, and fire growth rate that were used in the FDS simulation were estimated by taking into account all of the available information including the NIOSH report~\cite{NIOSH:Bowyer2}, post-incident pictures, and relevant literature. This resulted in a maximum specified source fire of approximately 20~MW in the basement and rear balcony of the structure. Based on the limited ventilation conditions in the basement, the FDS simulation results indicated a HRR of approximately 2~MW as a result of the limited supply of oxygen. After the rear basement windows failed, the HRR increased to approximately 32~MW before it reached a steady-state value of approximately 20~MW.
